---
title: "Choir Performance"
author: "MA676 A1 2024"
date: "2024-03-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
##I. Abstract
The success of a choir's choral performance has important implications for audiences, musicians, music institutions, etc. This report will mainly focus on predicting the sucess or failure of different choir performance. Specifically, the report will do analysis on Choir data with two method: EDA(Exploratory Data Analysis) and modeling. After reading this report, you will have a general idea how to predict the sucess or failure of choir and what's the most influential factor in predicting process.

##II. Introduction
This project will be focusing on the analysis on the relationship between sucess/failure and other factors on choirs. The client is doing a project for his research paper. This report focus will be on developing a regression prediction model so as to predict which factors are most likely to affect the performance of choirs, thereby providing robust support for our client's research. The whole project will consist of following parts "Abstract", "Introduction", "Objectives",  "Procedures", "Result","Discussion", "Reference" and "Appendix". 

    
##III. Objectives
The main obejecties include:
Whether there are significant differences between choir success scores and failure scores
Find out contributing factors to predict dimensional scores of school choir performance from conductor’s perspective


##IV. Procedures

###Data source

```{r echo=FALSE, message=FALSE, warning=FALSE}

#load packages
library(readr)
library(ggplot2)
library(tidyverse)
library(sqldf)
library(knitr)
library(lme4)
library(arm)
library(coefplot)
library(dplyr)
library(leaflet)
```
```

```{r message=FALSE, echo=FALSE, warning=FALSE}
#read in data and combine


####overview of Airbnb data:

####features of the dataset:

###EDA

